Generate Random Forest Models

### 2023-12-30 scripts/download_independent_variables.Rmd
reran soil grids

```{r download soil grids data eval = F}
library(terra)
library(sf)
library(gdalUtilities)

@@

vrts <- paste0(var, '/', var, '_', rep(depth, each = length(var)), 'cm_mean.vrt')

gdal_translate(
  paste0(sg_url, vrts),
               "./crop_roi_igh_r.vrt",
               of = "VRT", tr = c(250, 250),
               projwin = domain,
               projwin_srs = igh)

@@

#### salinity ####

@@

@@

filename <- paste0(pout, '/herb/', 'herb_.tif')
makeTiles(herbs, template, filename)
rm(herbs)

>>>>

```{r download soil grids data,  eval = F}
library(terra)
library(sf)
library(gdalUtilities)

@@

vrts <- paste0(var, '/', var, '_', rep(depth, each = length(var)), 'cm_mean.vrt') 
vrts

lapply(paste0(sg_url, vrts)[1:2], gdal_translate,  dst_dataset = "./crop_roi_igh_r.vrt",
               of = "VRT", tr = c(250, 250),
               projwin = domain,
               projwin_srs = igh)

gdal_translate(
  paste0(sg_url, vrts[1]),
               "./crop_roi_igh_r.vrt",
               of = "VRT", tr = c(250, 250),
               projwin = domain,
               projwin_srs = igh)

@@

@@

```{r create soil grids products, eval = F}

p2grids <- '/media/steppe/hdd/Geospatial_data/soilgrids'
prods <- list.files(p2grids)
tifs <- list.files(p2grids, recursive = TRUE)
tifs <- tifs[grepl('cm_mean', tifs)]

# tifs is input.

gridder <- function(x){

  product <- terra::vrt(file.path(p2grids, x))
  product <- terra::crop(product, terra::project(domain, terra::crs(product)))
  product <- terra::project(product, 'EPSG:5070')
  product <- resample(product, target_cells)

  out_root <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2/soilgrids'
  paths <- gsub('cm_.*', '', x)
  paths <- file.path(out_root, gsub('-', '_', paths))

  filename <- file.path(paths, basename(x))
  makeTiles(product, template, filename)

}

lapply(tifs, gridder)

p1 <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2/soilgrids/cfvo/cfvo_0_5'
cfvo5 <- vrt(file.path(p1, list.files(p1)))
plot(cfvo5)
```

@@

filename <- paste0(pout, '/herb/', 'herb_.tif')
makeTiles(herbs, template, filename)

### 2023-12-30 scripts/generate_RF_models.Rmd
replacing NA with biological 0 in 3 rasters
reran soil grids

```{r}
library(tidyverse)
library(terra)
source('functions.R')
```

@@

p <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('/chelsa$', paths) ]

# soils need to be uploaded seperately, e.g. by the depths
paths <- paths[! grepl('/soilgrids$', paths) ]

@@

layers <- lapply(files, vrt)
layers <- rast(layers)
names(layers) <- basename(paths)
plot(layers[[1:3]])

# gdgfgd10", "gdgfgd5, "swe"
# both need NA replaced with 0, if terrestrial...

layers[[19]] <- subst(layers[[19]], NA, 0)
layers[[19]] <- mask(layers[[19]], layers[[1]]) # remove the ocean/set extent again

plot(layers[[22]])

```

@@

abvi_test <- terra::extract(layers, abvi)
```

>>>>

```{r}
library(tidyverse)
library(terra)
library(sf)
source('functions.R')
```

@@

p <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('chelsa$|veg_cover$|soilgrids$|sand$|silt$|clay$|phh2o$|cfvo$', paths) ]

@@

layers <- lapply(files, vrt)
layers <- rast(layers)
names(layers) <- basename(paths)
```

@@

abvi_test <- terra::extract(layers, abvi)

# identify a single component of each pair for usage. 
```
